---
title: "Introduction to Tidy Finance"
author: "Jens Pedersen"
date: '2023-01-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2 - Introduction to Tidy Finance

In this R Markdown document, you can find my solutions to the exercises for chapter 2 in the course Advanced Empirical Finance.

## 2.1 - Working with stock market data

```{r Question 1}
library(tidyquant)
library(tidyverse)
library(ggplot2)
AAPL <- tq_get("AAPL",
  get = "stock.prices",
  from = "2000-01-01",
  to = "2021-12-31")

AAPL |>
  ggplot(aes(x = date, y = adjusted)) +
  geom_line() +
  labs(
    x = NULL,
    y = NULL,
    title = "Apple stock prices between beginning of 2000 and end of 2021"
  )
```

## 2.2 - Working with stock market data

Now, we would like to analyse the daily net returns, which are defined as r_t = p_t/p\_{t-1}-1.

```{r Question 2}
returns <- AAPL |>
  arrange(date) |>
  mutate(ret = adjusted / lag(adjusted) - 1) |>
  select(symbol, date, ret)
returns
```

The above tibble consists of three columns, where the last contains the daily returns (ret). Naturally, the first row in column (ret) is NA. The arrange() provides a convenient way to order observations in the correct ascending rate. We could write arrange(desc(date)) to order observations by descending dates.

Furthermore, we would like to remove missing values.

```{r Question 2}
returns <- returns |>
  drop_na(ret)
```

Next up, we visualise the distribution of daily returns in a histogram. Just for convenience, we multiply the returns by 100 to get returns in per cent for the visualisation.

```{r Question 2}
quantile_05 <- quantile(returns |> pull(ret) * 100, probs = 0.05)

returns |>
  ggplot(aes(x = ret * 100)) +
  geom_histogram(bins = 100) +
  geom_vline(aes(xintercept = quantile_05),
    linetype = "dashed"
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Distribution of daily Apple stock returns in percent"
  )
```

## 2.3 - Compute summary statistics

```{r Question 3}

returns |>
  mutate(ret = ret * 100) |>
  summarize(across(
    ret,
    list(
      daily_mean = mean,
      daily_sd = sd,
      daily_min = min,
      daily_max = max
    )
  ))

```

Furthermore, we can compute the summary statistics for each individual year with the following chunk.

```{r Question 3}
returns |>
  mutate(ret = ret * 100) |>
  group_by(year = year(date)) |>
  summarize(across(
    ret,
    list(
      daily_mean = mean,
      daily_sd = sd,
      daily_min = min,
      daily_max = max
    ),
    .names = "{.fn}"
  )) |>
  print(n = Inf)

```

## 2.2 - Scaling up the analysis

For this part, we would like to scale up our previous analysis to include "MMM" and "BA". This code is homemade, please do ask, if you have any questions.

```{r Question 1}

stock_tickers <- c("AAPL", "MMM", "BA")
placeholder_tickers <- list()

for (i in stock_tickers) {
  placeholder_tickers[[i]] <- tq_get(i,
  get = "stock.prices",
  from = "2000-01-01",
  to = "2021-12-31")
}
# This library is necessary in order to execute plot_grid(plotlist=plots), which stack plots in order of the above for loop.
library(cowplot)
plots <- list()

for (i in stock_tickers) {
  plots[[i]] <- local({
    i <- i
    p <- as.data.frame(placeholder_tickers[[i]]) |>
      ggplot(aes(x = date, y = adjusted)) +
      geom_line() +
      labs(
      x = NULL,
      y = NULL,
    title =  paste0(i," stock prices between beginning of 2000 and end of 2021")
  )
  })
}

plot_grid(plotlist=plots)

```

Now, we would like to compute the returns, but for the vector.

```{r Question 1 cntd.}

returns_vector <- list()

for (i in stock_tickers) {
returns_vector[[i]] <- placeholder_tickers[[i]] |>
  arrange(date) |>
  mutate(ret = adjusted / lag(adjusted) - 1) |>
  select(symbol, date, ret) |>
  filter(!is.na(ret)) # this line of code is added to remove NAs
}

returns

```

Next, we compute the histogram for each stock in the vector.

```{r Question 1 cntd.}

histogram <- list()

for (i in stock_tickers) {
  histogram[[i]] <- local({
    i <- i
    h <- as.data.frame(returns_vector[[i]]) |>
          ggplot(aes(x = ret * 100)) +
          geom_histogram(bins = 100) +
          geom_vline(aes(xintercept = quantile(returns_vector[[i]] |> pull(ret) * 100, probs = 0.05)),
          linetype = "dashed") +
          labs(
                x = NULL,
                y = NULL,
                title =  paste0(i, " Distribution of daily Apple stock returns in percent"))
  })
}

plot_grid(plotlist=histogram)
```

Now, we compute the summary statistics for the vector.

```{r Question 1 cntd.}

for (i in stock_tickers) {
returns_vector[[i]] |>
  mutate(ret = ret * 100) |>
  summarize(across(
    ret,
    list(
      daily_mean = mean,
      daily_sd = sd,
      daily_min = min,
      daily_max = max
    )
  )) 
}
returns_vector

```

```{r Question 1 cntd.}
for (i in stock_tickers) {
returns_vector[[i]] |>
  mutate(ret = ret * 100) |>
  group_by(year = year(date)) |>
  summarize(across(
    ret,
    list(
      daily_mean = mean,
      daily_sd = sd,
      daily_min = min,
      daily_max = max
    ),
    .names = "{.fn}"
  )) |>
  print(n = Inf)

  }
```

Now, we will call the constituents of the Dow Jones index.

```{r Question 1 cntd.}
ticker <- tq_index("DOW")
ticker

index_prices <- tq_get(ticker,
  get = "stock.prices",
  from = "2000-01-01",
  to = "2021-12-31"
)

```

Plotting

```{r Question 1 cntd.}
index_prices |>
  ggplot(aes(
    x = date,
    y = adjusted,
    color = symbol
  )) +
  geom_line() +
  labs(
    x = NULL,
    y = NULL,
    color = NULL,
    title = "Stock prices of DOW index constituents"
  ) +
  theme(legend.position = "none")
```

Summarizing

```{r Question 1 cntd.}
all_returns <- index_prices |>
  group_by(symbol) |>
  mutate(ret = adjusted / lag(adjusted) - 1) |>
  select(symbol, date, ret) |>
  drop_na(ret)

all_returns |>
  mutate(ret = ret * 100) |>
  group_by(symbol) |>
  summarize(across(
    ret,
    list(
      daily_mean = mean,
      daily_sd = sd,
      daily_min = min,
      daily_max = max
    ),
    .names = "{.fn}"
  )) |>
  print(n = Inf)

```
Now, we would like to answer the following question: Are days with high aggregate trading volume often followed by high aggregate trading volume days?
```{r Question 2}
volume <- index_prices |>
  group_by(date) |>
  summarize(volume = sum(volume * close / 1e9))

volume |>
  ggplot(aes(x = date, y = volume)) +
  geom_line() +
  labs(
    x = NULL, y = NULL,
    title = "Aggregate daily trading volume of DOW index constitutens"
  )
```
From the above figure, we clearly see that there has been an upward trend in the daily trading volume, which has been particularly high after the COVID-19 outbreak. We can also illustrate the persistence of trading volume by plotting the volume at time t against t-1.
```{r Question 2 cntd.}
volume |>
  ggplot(aes(x = lag(volume), y = volume)) +
  geom_point() +
  geom_abline(aes(intercept = 0, slope = 1),
    linetype = "dashed"
  ) +
  labs(
    x = "Previous day aggregate trading volume",
    y = "Aggregate trading volume",
    title = "Persistence in daily trading volume of DOW index constituents"
  )

```
From the above, it is clear that days with high trading volume are often followed by similarly high trading volume days.
