---
title: "Portfolio choice problems"
author: "Jens Pedersen"
date: '2023-02-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2 - Introduction to Tidy Finance

In this R Markdown document, you can find my solutions to the exercises for chapter 2 in the course Advanced Empirical Finance.

## 2.3 - Portfolio choice problems

First, we will compute the monthly returns from all adjusted daily Dow Jones constituent prices

```{r Question 1}
library(tidyquant)
library(tidyverse)
library(ggplot2)

ticker <- tq_index("DOW")

index_prices <- tq_get(ticker,
  get = "stock.prices",
  from = "2000-01-01",
  to = "2021-12-31"
)

index_prices <- index_prices |>
  group_by(symbol) |>
  mutate(n = n()) |>
  ungroup() |>
  filter(n == max(n)) |>
  select(-n)


returns <- index_prices |>
  mutate(month = floor_date(date, "month")) |>
  group_by(symbol, month) |>
  summarize(price = last(adjusted), .groups = "drop_last") |>
  mutate(ret = price / lag(price) - 1) |>
  drop_na(ret) |>
  select(-price)

returns
```

Next, we compute the vector of historical average returns and sample variance-covariance matrix.

```{r Question 2}
returns_matrix <- returns |>
  pivot_wider(
    names_from = symbol,
    values_from = ret
  ) |>
  select(-month)

Sigma <- cov(returns_matrix)
mu <- colMeans(returns_matrix)
```

Now, we compute the minimum variance portfolio weights and the portfolio volatility and average returns.

```{r Question 3}
N <- ncol(returns_matrix)
iota <- rep(1, N)
mvp_weights <- solve(Sigma) %*% iota
mvp_weights <- mvp_weights / sum(mvp_weights)

tibble(
  average_ret = as.numeric(t(mvp_weights) %*% mu),
  volatility = as.numeric(sqrt(t(mvp_weights) %*% Sigma %*% mvp_weights))
)
```

Note that the command solve(A,b) solves the equation Ax=b, if b is not provided as in the snippet above defaults to the identity matrix such that solve(Sigma) is delivers Sigma\^-1.

Lastly, we want to visualise the mean-variance efficient frontier.

```{r Question 4}
mu_bar <- 3 * t(mvp_weights) %*% mu

C <- as.numeric(t(iota) %*% solve(Sigma) %*% iota)
D <- as.numeric(t(iota) %*% solve(Sigma) %*% mu)
E <- as.numeric(t(mu) %*% solve(Sigma) %*% mu)

lambda_tilde <- as.numeric(2 * (mu_bar - D / C) / (E - D^2 / C))
efp_weights <- mvp_weights +
  lambda_tilde / 2 * (solve(Sigma) %*% mu - D * mvp_weights)

c <- seq(from = -0.4, to = 1.9, by = 0.01)
res <- tibble(
  c = c,
  mu = NA,
  sd = NA
)

for (i in seq_along(c)) {
  w <- (1 - c[i]) * mvp_weights + (c[i]) * efp_weights
  res$mu[i] <- 12 * 100 * t(w) %*% mu   
  res$sd[i] <- 100 * sqrt(12) * sqrt(t(w) %*% Sigma %*% w)
}

res |>
  ggplot(aes(x = sd, y = mu)) +
  geom_point() +
  geom_point(
    data = res |> filter(c %in% c(0, 1)),
    size = 4
  ) +
  geom_point(
    data = tibble(
      mu = 12 * 100 * mu,       
      sd = 100 * sqrt(12) * sqrt(diag(Sigma))
    ),
    aes(y = mu, x = sd), size = 1
  ) +
  labs(
    x = "Annualized standard deviation (in percent)",
    y = "Annualized expected return (in percent)",
    title = "Efficient frontier for DOW index constituents"
  )
```
